\replaced{Fast Fourier transformation of electromagnetic data for expensive kernels}{Fast time-domain electromagnetic modelling in the frequency domain}

### 2020-11-26 ManuscriptTime.tex
Within revision

% \documentclass[onecolumn,extra,referee,camera]{gji}
% \documentclass[onecolumn,extra,camera]{gji}
\documentclass[twocolumn,extra,camera]{gji}
\usepackage{timet}

% Additional packages
% \usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{ifthen}
\usepackage[draft, addedmarkup=bf]{changes}
% \usepackage[final]{changes}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}

@@

% Figures
% To print widths put this in text:
% \the\textwidth
% \the\columnwidth
%
% Widths:   textwidth  columnwidth  tablewidth
%       1-c   504pt        252pt       504pt
%       2-c   504pt        240pt       240pt
%       ref   468pt        234pt       468pt
%
\ifthenelse{\boolean{GJ@referee}}{%
  \newcommand{\cwidth}{234pt}%
  \newcommand{\fwidth}{468pt}%
  \newcommand{\twidth}{468pt}}{%
  \newcommand{\fwidth}{504pt}%
  \ifthenelse{\boolean{@twocolumn}}{%
    \newcommand{\cwidth}{240pt}%
    \newcommand{\twidth}{240pt}}{%
    \newcommand{\cwidth}{252pt}%
    \newcommand{\twidth}{504pt}%
  }
}
\graphicspath{{./figures/}}

% Change default colour of «changes» package
\definecolor{myred}{rgb}{.8,0,0}
\definecolor{mygreen}{rgb}{0,.5,0}
% \setaddedmarkup{\color{myred}{#1}}
\setaddedmarkup{\color{myred}{#1}}
\setdeletedmarkup{\color{mygreen}{\sout{#1}}}

@@

% Remove once done
\newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
\newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
\newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
\newcommand{\dul}[1]{\underline{\underline{#1}}}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\rmk}[1]{{\color{red}{#1}}}

\title[Time-domain CSEM]{Fast time-domain electromagnetic modelling in the frequency domain}

\author[D. Werthmüller \emph{et al.}]
  {\Large Dieter Werthmüller$^1$,        % orcid: 0000-0002-8575-2484
   Wim A. Mulder$^{1,2}$ and          % orcid: 0000-0001-7020-9297
   Evert C. Slob$^1$              % orcid: 0000-0002-4529-1134
   \\
   \footnotesize
  $^1$ TU Delft, Building 23, Stevinweg 1 / PO-box 5048, 2628 CN Delft (NL).
  E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org}\\[-.3em]
   \footnotesize
  $^2$ Shell Global Solutions International BV, Grasweg 31, 1031 HW Amsterdam(NL)
  }

@@

\begin{document}

@@

\begin{summary}
%
Modelling three-dimensional time-domain controlled-source electromagnetic data
with frequency-domain solvers is an alternative to time-domain solvers. It
requires the computation of many frequencies if standard Fourier transforms are
used, which can make it prohibitively expensive in comparison with a
time-domain computation. The speed of time-domain modelling with
frequency-domain computation is defined by three key points: solver, method and
implementation of the Fourier transformation, and gridding.
%
The faster the frequency-domain solver, the faster time-domain modelling will
be. It is important that the solver is robust over a wide range of frequencies.
A three-dimensional, iterative, frequency-domain multigrid solver satisfies
this requirement, but any other robust solver will do.
%
The method should require as few frequencies as possible while remaining
robust. As the frequency range spans many orders of magnitude, the required
frequencies are ideally equally spaced on a logarithmic scale. The fast
frequency-to-time domain method uses either the digital linear filter method
or the logarithmic fast Fourier transform together with a careful selection of
frequencies and interpolation. This methodology requires typically 15 to 20
frequencies to cover a wide range of offsets in one-dimensional test models.
%
The gridding should be frequency-dependent, which is accomplished by making it
a function of skin depth. Optimising for the least number of required cells for
a given frequency should be combined with optimising for computational speed.
%
Looking carefully at these three \added{key} points results in much smaller
computation times with speedup factors of ten or more over previous results. A
frequency-domain computation of time-domain data can therefore be competitive
to time-domain solvers if the required frequencies and the corresponding grids
are carefully chosen.
%
\end{summary}

@@

@@

For the numerical computations we use the open-source (Apache License 2.0) code
\emg3d \citep{JOSS.19.Werthmuller}, a multigrid solver based on
\cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov subspace
solvers or as a solver on its own. The multigrid approach works fine for the
diffusive approximation of Maxwell's equation, which assumes that
$\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
(F/m). The remaining system to solve in the frequency domain is then given by
the second-order differential equation of the electric field,
%
\begin{equation}
    \mathrm{i}\omega\sigma \mathbf{E} +
    \nabla \times \mu^{-1} \nabla \times \mathbf{E}
    = -\mathrm{i}\omega\mathbf{J}_\mathrm{s} \, ,
  \label{eq:maxwell}
\end{equation}
%
where $\mathbf{E}$ is the electric field (V/m) and $\mathbf{J}_\mathrm{s}$ the
current source (A/m$^2$); time dependence is $\exp(\mathrm{i}\omega t)$. The
standard multigrid approach fails for severe stretching or strong anisotropy,
for which \deleted{\emg3d has implemented }known improvements such as
line-relaxation and semicoarsening \citep{ECCFD.06.Jonsthovel} \added{are
implemented}, with a non-standard Cholesky decomposition to speed up the
computation \citep{GEO.08.Mulder}. One of the big advantages of the multigrid
method is that it scales linearly (optimal) with the grid size in both CPU and
RAM usage \citep{B.Springer.20.Mulder}. This makes it feasible to run even big
models on standard computers, without the need for big clusters. All the
examples in this article are run on a laptop with an i7-6600U\,CPU\,@\,2.6\,GHz
(x4) and 16\,GB of memory, using Ubuntu 18.04 and Python 3.7. For the
semi-analytical computations of layered models we use the open-source (Apache
License 2.0) code \empymod \citep{GEO.17.Werthmuller}.

@@

An important factor in terms of speed and accuracy for time-domain responses
obtained from frequency-domain computations is the selection of the required
frequencies. The fewer frequencies required, the quicker we obtain the
time-domain result. We decided to use a regular spacing of frequencies on a
log-scale, rather than an adaptive scheme. This approach is favourable if a
wide range of offsets is needed, as the required frequencies change with offset
and an adaptive frequency selection is therefore often tailored to a single
offset. Also, it allows for straightforward parallelization over frequencies,
which is not completely possible with an adaptive scheme.

For the actual transform we use either the digital linear filter (DLF) method
or the logarithmic fast Fourier transform (FFTLog). The DLF method was
introduced to geophysics by \cite{GP.71.Ghosh}, and is arguably the most common
method in EM geophysics for both its simplicity and its speed. It is
implemented for the Hankel and Fourier transforms in most EM modelling codes,
e.g., \cite{GEO.09.Key}. A simple tool to design digital linear filters was
recently presented by \cite{GEO.19.Werthmuller}, together with a comprehensive
overview of the history and development of DLF in geophysics. FFTLog,
introduced by \cite{RAS.00.Hamilton}, is another transform algorithm which
proved to be powerful for the frequency-to-time domain transformation of EM
responses, e.g., \cite{INT.14.Werthmuller}. In our tests they are both about
equal in speed and accuracy. DLF requires a wider range and many more
frequencies than the FFTLog. Both methods share some important characteristics
in comparison with the standard FFT: the required input frequencies are equally
spaced on a logarithmic scale, and they only require either the real or the
imaginary part of the frequency-domain response. We can take advantage of that
by using only the imaginary part of the frequency-domain response. The
imaginary part goes to zero when the frequency goes to zero or to infinity,
with the advantage that knowing the endpoints makes it possible to convert the
extrapolation of missing frequencies into interpolation.

There are three key-parameters that have to be defined: Minimum and maximum
frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies per decade
($n_\mr{dec}^f$). Trial-and-error with a 3D code is very time-intensive.
However, a simplified, layered model for the required survey setup and a fast
1D modeller makes it possible to estimate these parameters easily. To this end,
we created a graphical, interactive tool using \empymod, as shown in
Figure~\ref{fig:GUI-FFTLog-lin}.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{GUI-FFTLog-lin}
  \caption{Interactive frequency selection for a user-provided layered model
    (the shown model parameters are described in the text). This example shows
    the impulse response at an offset of 5\,km, for which it uses FFTLog from
    0.001\,Hz to 10\,Hz with five frequencies per decade. Note that the error
    is clipped for values smaller than 0.01\,\% and bigger than 100\,\%.}
  \label{fig:GUI-FFTLog-lin}
\end{figure}
%

The example model is a marine scenario with 1\,km water depth of resistivity
$\rho=0.3\,\Omega$\,m ($\rho = \sigma^{-1}$) and a 100\,m thick target of
100\,$\Omega$\,m at 1\,km below the seafloor in a background of 1\,$\Omega$\,m.
The source is 50\,m above the seafloor, the receiver is on the seafloor, and
the response is the inline $x$-directed $E$-field. The left subplot shows the
imaginary part in the frequency domain and the right plot the corresponding
impulse response in the time domain. The red lines are the
\replaced{semi-analytical responses}{precise results obtained from \empymod}.
The blue circles indicate the actually computed responses and the black dots
the frequencies which are interpolated or set to zero. The resulting
time-domain response has a relative error of less than 1\,\% everywhere except
for very early times.

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\begin{figure*}
  \centering
  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-FFTLog-log}}%
  \label{fig:GUI-FFTLog-log}}\hfill
  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-DLF-lin}}%
  \label{fig:GUI-DLF-lin}}\\
  \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
  scale. (b) Using DLF instead of FFTLog, but computing the same
  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
  response has comparable accuracy (compare the error to the error in Figure
  \ref{fig:GUI-FFTLog-lin}).}
  \label{fig:GUI-log-lin}
\end{figure*}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. However, with this frequency-selection we can
model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
where the above parameters were used to model the responses at offsets
$r=1.5,3,6$, and 12\,km.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{multi-offset}
  \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
    $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
    \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
    responses, the black dots are the responses which are set to zero or
    interpolated. The black curves are the analytical responses.}
  \label{fig:multi-offset}
\end{figure}
%

The shortest offset defines the highest required frequency, and the largest
offset the lowest required frequency. So the more one can restrict the
necessary offset range, the \replaced{fewer}{less} frequencies are needed.
Another important factor is how to interpolate and extrapolate from the
computed frequencies to the frequencies required for the Fourier transform. For
the FFTLog only extrapolation for higher and lower frequencies is required. The
EM response becomes highly oscillatory for high frequencies, which makes it
very hard to extrapolate the response to frequencies $f>f_\mathrm{max}$.
However, if $f_\mathrm{max}$ is chosen judiciously, the importance of higher
frequencies for the Fourier transform can be neglected and we can set those
responses to zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be
changed to an interpolation by assuming a zero imaginary response at zero
frequency, and we then use PCHIP to interpolate the missing frequencies (as we
work on a logarithmic scale we cannot choose 0\,Hz, but instead take
$10^{-100}$\,Hz). In the case of the DLF method we also have to interpolate in
between the computed frequencies, for which we found it better to use a cubic
spline. As can be seen from Figures \ref{fig:GUI-FFTLog-lin} and
\ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF with the same actually
computed frequencies results in very similar responses.

@@

The main difference with \cite{GEO.08.Mulder} is that their adaptive gridding
searches for the optimal stretching factor $\alpha$ fulfilling certain
criteria, for a fixed number of cells. Our adaptive gridding, on the other
hand, searches for the smallest number of cells that still fulfil the given
criteria. The number of cells becomes therefore also a function of frequency.
To go from the model grid to the computational grid, we use the
volume-averaging technique on logarithmic resistivities, as used in
\cite{GEO.07.Plessix}. While this technique ensures that the total resistivity
in the subsurface remains the same, it does not consider effective-medium
theory \citep{GEO.03.Davydycheva}, for instance, the apparent anisotropy from a
stack of finely layered formations of varying resistivity.

>>>>

\documentclass[onecolumn,extra,referee,camera]{gji}
% \documentclass[onecolumn,extra,camera]{gji}
% \documentclass[twocolumn,extra,camera]{gji}
\usepackage{timet}

% Additional packages
% \usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage[position=top]{subfig}
\usepackage{ifthen}
\usepackage[draft, addedmarkup=bf]{changes}
% \usepackage[final]{changes}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}

@@

% Figures
% To print widths put this in text:
% \the\textwidth
% \the\columnwidth
%
% Widths:   textwidth  columnwidth  tablewidth
%       1-c   504pt        252pt       504pt
%       2-c   504pt        240pt       240pt
%       ref   468pt        234pt       468pt
%
\newlength{\fwidth}
\newlength{\cwidth}
\newlength{\twidth}
\ifthenelse{\boolean{GJ@referee}}{%
  \setlength{\cwidth}{234pt}%
  \setlength{\fwidth}{468pt}%
  \setlength{\twidth}{468pt}}{%
  \setlength{\fwidth}{504pt}%
  \ifthenelse{\boolean{@twocolumn}}{%
    \setlength{\cwidth}{240pt}%
    \setlength{\twidth}{240pt}}{%
    \setlength{\cwidth}{252pt}%
    \setlength{\twidth}{504pt}%
  }
}
\graphicspath{{./figures/}}

% Change default colour of «changes» package
\definecolor{alizarin}{rgb}{0.82, 0.1, 0.26}
\definecolor{mygreen}{rgb}{0, .5, 0}
\setaddedmarkup{\color{mygreen}{\uline{#1}}}
\setdeletedmarkup{\color{alizarin}{\uwave{#1}}}

@@

% Remove once done
\newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
\newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
\newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
\newcommand{\dul}[1]{\underline{\underline{#1}}}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\rmk}[1]{{\color{red}{#1}}}
\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}

\title[\replaced{Transform for expensive EM kernels}{Time-domain CSEM}]{
  \replaced{Fast Fourier transformation of electromagnetic data for expensive kernels}{Fast time-domain electromagnetic modelling in the frequency domain}}

\author[D. Werthmüller \emph{et al.}]
  {\Large
   Dieter Werthmüller$^1$,     % orcid: 0000-0002-8575-2484
   Wim A. Mulder$^{1,2}$ and   % orcid: 0000-0001-7020-9297
   Evert C. Slob$^1$           % orcid: 0000-0002-4529-1134
   \\
   {\footnotesize
    $^1$ TU Delft, Building 23, Stevinweg 1 / PO-box 5048, 2628 CN Delft (NL).
    E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org}
   }\\[-.3em]
   {\footnotesize
    $^2$ Shell Global Solutions International BV, Grasweg 31, 1031 HW
    Amsterdam(NL)
   }
  }

@@

\begin{document}
\large %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%

@@

\begin{summary}
%
Modelling three-dimensional time-domain controlled-source electromagnetic data
with frequency-domain solvers is an alternative to time-domain solvers. It
requires the computation of many frequencies if standard Fourier transforms are
used, which can make it prohibitively expensive in comparison with a
time-domain computation. The speed of time-domain modelling with
frequency-domain computation is defined by three key points: solver, method and
implementation of the Fourier transformation, and gridding.
%
The faster the frequency-domain solver, the faster time-domain modelling will
be. It is important that the solver is robust over a wide range of frequencies.
A three-dimensional, iterative, frequency-domain multigrid solver satisfies
this requirement, but any other robust solver will do.
%
The method should require as few frequencies as possible while remaining
robust. As the frequency range spans many orders of magnitude, the required
frequencies are ideally equally spaced on a logarithmic scale. The fast
frequency-to-time domain method uses either the digital linear filter method
or the logarithmic fast Fourier transform together with a careful selection of
frequencies and interpolation. This methodology requires typically 15 to 20
frequencies to cover a wide range of offsets\deleted{in one-dimensional test
models}.
%
The gridding should be frequency-dependent, which is accomplished by making it
a function of skin depth. Optimising for the least number of required cells for
a given frequency should be combined with optimising for computational speed.
%
Looking carefully at these \deleted{three} points results in much smaller
computation times with speedup factors of ten or more over previous results. A
frequency-domain computation of time-domain data can therefore be competitive
to time-domain solvers if the required frequencies and the corresponding grids
are carefully chosen.
%
\end{summary}

@@

{\color{alizarin}
\verb!v v v REMOVE COMMENTS v v v REMOVE COMMENTS v v v REMOVE COMMENTS v v v!
}
{\color{mygreen}
\begin{itemize}
  \item Too generic; more references
  \item what are the methodological advantages with respect to other
    approaches.
\end{itemize}

Redo intro: Focus on transform, not on frequency- or time-domain EM.

\begin{itemize}
  \item Wim: Include perhaps (one of) explicitly, instead of implicitly:
    Druskin, V. L., and L. A. Knizhnerman, 1994, Spectral approach to solving
    three-dimensional Maxwell’s diffusion equations in the time and frequency
    domains, Radio Science, 29, 937--953. doi: 10.1029/94RS00747 Druskin, V.
    L., L. A. Knizhnerman, and P. Lee, 1999, New spectral Lanczos decomposition
    method for induction modeling in arbitrary 3D geometry, Geophysics, 64,
    701--706. doi: 10.1190/1.1444579.
  \item Evert: I also think we should respond a bit more accommodating the
    reviewer's request. Mulder 2008 is quite old now and papers were published
    after that dealing with solutions directly in time domain. We can write
    that in this paper we include only papers that were not reference in Mulder
    et al. (2008) and refer to that for earlier papers. New are Mittet 2010
    (FDTD of waves and correspondence principle, GPY), Um et al. 2010 (Finite
    element, SEG abstract) and Zhou et al. 2020 TGRS
    (10.1109/TGRS.2020.2979798). In terms of number of references, you can
    decide to not worry to much about that, or remove some others. You have
    several references to works on the differences between measuring/analysing
    data in time or frequency domain. That could go out, e.g.
  \item Evert: i think you can group separate points under the recommendation
    that comes from his first concern. I understand his recommendation as:
    "Write an introduction that is more focused on time domain CSEM modeling."
    That should be done in a comprehensive way and include direct time domain
    methods such as Krylov subspace, Simpeg, FDTD of waves and use
    correspondence principle. It can be interesting to also include more
    arbitrary source time functions and not only step switch on or off and
    impulse response. For the filter approach, in principle, all FD codes are
    suitable to be used. His second concern phrased as a recommendation: "Show
    results from a more complex model." after that the specific points come.
  \item Evert: I think that Commer wanted to refer to Druskin and Knihznerman
    (i am sure spelling is not correct) in radio science 1994, which is most
    likely the older version of the code he mentions. We refer to that 1994
    paper in Mulder 2008.
  \item Evert: The main issue is that the Druskin work it is what Commer uses.
    This is the principal issue. This comes back at point 9. Perhaps now you
    can include it in the introduction as a motivation to work on this topic.
  \item Evert: So let us make an explicit sentence in the introduction to
    clarify this. We also may need to change the text in the introduction to
    avoid making our life difficult by selecting some papers while not
    selecting others. This can be done by only selecting review papers on the
    topic and write that that is our choice here.
  \item Wim: Is that different from doi 10.1190/1.1801936 (mentioned in
    Mulder,Wirianto,Slob)? Then also include Haber from that paper, or just all
    references of section 2 (complexity) in that paper?
  \item Evert: This all belongs to one concern: The introduction is too general
    and insufficiently addresses the actual topic of obtaining time domain data
    by CSEM modeling. That does not mean you need to repeat everything. You
    could start by writing that we will only discuss new developments since
    2008, and refer to Mulder 2008 for older works that relate to the topic of
    time domain CSEM data modeling. I don't think Commer actually wants an
    extensive review of the literature and include his work. He writes that the
    introduction should be more focused on the topic of the paper.
  \item Evert: i have no problem with mentioning in the paper that our idea
    grid coarsening can be used in the method of Commer and Newman (2005). In
    that case no further response is necessary, we tick it off by saying, we
    refer to Commer and Newman (2005) on grid-coarsening. i would limit
    references to Commer to this paper then. I don't think we say anywhere that
    our method is the best in any sense. It seems this reviewer is in
    competition. I don't think we should be in competition. We work for the
    public domain. We put forward our ideas. We leave to the readers to decide
    which options they prefer. We should not be forced need to sell our method.
    Of course, my personal opinion does not change the reality that the AE has
    written that he wants to see the majority of the reviewers comments
    implemented in our revision. We should accept that.
  \item Evert: If we judge this is part of the majorty to be implemented as the
    Associate Editor says, it means the AE agrees with Commer and we should do
    it. No need to add references. In the introduction, just add a summary of
    the order analysis given in Mulder 2008 where some order analysis was
    done.
  \item References provided by Commer:
    \begin{itemize}
      \item Commer \& Newman, 2005, An accelerated time domain finite
        difference simulation scheme for three-dimensional transient
        electromagnetic modeling using geometric multigrid concepts, Radio
        Science, 41, RS3007.
      \item Commer et al., 2006, New results on the resistivity structure of
        Merapi Volcano(Indonesia), derived from three-dimensional restricted
        inversion of long-offset transient electromagnetic data, GJI.
      \item Druskin \& Zaslavsky, 2011, Solution of 3D time-domain
        electromagnetic problems using optimal subspace projection,
        Geophysics.
      \item Weiland, T., 1996, Time domain EM field computation with finite
        difference methods, Int. J. Num. Mod., 9, 295-319.
      \item Yuji Mitsuhata, Toshihiro Uchida, Yutaka Murakami, Hiroshi Amano,
        The Fourier transform of controlled-source time-domain electromagnetic
        data by smooth spectrum inversion, Geophysical Journal International,
        144, January 2001, 123-135.
    \end{itemize}
\end{itemize}

}
{\color{alizarin}
\verb!^ ^ ^ REMOVE COMMENTS ^ ^ ^ REMOVE COMMENTS ^ ^ ^ REMOVE COMMENTS ^ ^ ^!
}

@@

For the numerical computations we use the open-source (Apache License 2.0) code
\emg3d \citep{JOSS.19.Werthmuller}, a multigrid solver based on
\cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov subspace
solvers or as a solver on its own. The multigrid approach works fine for the
diffusive approximation of Maxwell's equation, which assumes that
$\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
(F/m). The remaining system to solve in the frequency domain is then given by
the second-order differential equation of the electric field,
%
\begin{equation}
    \mathrm{i}\omega\sigma \mathbf{E} +
    \nabla \times \mu^{-1} \nabla \times \mathbf{E}
    = -\mathrm{i}\omega\mathbf{J}_\mathrm{s} \, ,
  \label{eq:maxwell}
\end{equation}
%
where $\mathbf{E}$ is the electric field (V/m) and $\mathbf{J}_\mathrm{s}$ the
current source (A/m$^2$); time dependence is $\exp(\mathrm{i}\omega t)$. The
standard multigrid approach fails for severe stretching or strong anisotropy,
for which \deleted{\emg3d has implemented }known improvements such as
line-relaxation and semicoarsening \citep{ECCFD.06.Jonsthovel} \added{are
implemented}, with a non-standard Cholesky decomposition to speed up the
computation \citep{GEO.08.Mulder}. One of the big advantages of the multigrid
method is that it scales linearly (optimal) with the grid size in both CPU and
RAM usage \citep{B.Springer.20.Mulder}. This makes it feasible to run even big
models on standard computers, without the need for big clusters. All the
examples in this article are run on a laptop with an i7-6600U\,CPU\,@\,2.6\,GHz
(x4) and 16\,GB of memory, using Ubuntu \replaced{20.04}{18.04} and Python
\replaced{3.8}{3.7}. \added{All examples were run using a single thread, but
parallelization over frequencies is straight forward.} For the semi-analytical
computations of layered models we use the open-source (Apache License 2.0) code
\empymod \citep{GEO.17.Werthmuller}.

@@

An important factor in terms of speed and accuracy for time-domain responses
obtained from frequency-domain computations is the selection of the required
frequencies. The fewer frequencies required, the quicker we obtain the
time-domain result. \added{The same applies for the reverse operation, the
fewer times required, the quicker we obtain frequency-domain results. Even
though we limit our analysis here to frequency-to-time transformation, most of
the arguments are reversible and applicable to time-to-frequency
transformations.} We decided to use a regular spacing of frequencies on a
log-scale, rather than an adaptive scheme. This approach is favourable if a
wide range of offsets is needed, as the required frequencies change with offset
and an adaptive frequency selection is therefore often tailored to a single
offset. Also, it allows for straightforward parallelization over frequencies,
which is not completely possible with an adaptive scheme.

For the actual transform we use either the digital linear filter (DLF) method
or the logarithmic fast Fourier transform (FFTLog). The DLF method was
introduced to geophysics by \cite{GP.71.Ghosh}, and is arguably the most common
method in EM geophysics for both its simplicity and its speed. It is
implemented for the Hankel and Fourier transforms in most EM modelling codes,
e.g., \cite{GEO.09.Key}. A simple tool to design digital linear filters was
recently presented by \cite{GEO.19.Werthmuller}, together with a comprehensive
overview of the history and development of DLF in geophysics. FFTLog,
introduced by \cite{RAS.00.Hamilton}, is another transform algorithm which
proved to be powerful for the frequency-to-time domain transformation of EM
responses, e.g., \cite{INT.14.Werthmuller}. In our tests they are both about
equal in speed and accuracy. DLF requires a wider range and many more
frequencies than the FFTLog. Both methods share some important characteristics
in comparison with the standard FFT: the required input frequencies are equally
spaced on a logarithmic scale\added{ (natural logarithm in the case of the DLF
and decimal logarithm in the case of the FFTLog)}, and they only require either
the real or the imaginary part of the frequency-domain response. We can take
advantage of that by using only the imaginary part of the frequency-domain
response. The imaginary part goes to zero when the frequency goes to zero or to
infinity, with the advantage that knowing the endpoints makes it possible to
convert the extrapolation of missing frequencies into \added{an }interpolation.

There are three key-parameters that have to be defined: Minimum and maximum
frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies per decade
($n_\mr{dec}^f$). Trial-and-error with a 3D code is very time-intensive.
However, a simplified, layered model for the required survey setup and a fast
1D modeller makes it possible to estimate these parameters easily. To this end,
we created a graphical, interactive tool using \empymod, as shown in
Figure~\ref{fig:GUI-FFTLog-lin}.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{01-GUI-FFTLog-lin}
  \caption{Interactive frequency selection for a user-provided layered model
    (the shown model parameters are described in the text). This example shows
    the impulse response at an offset of 5\,km, for which it uses FFTLog from
    0.001\,Hz to 10\,Hz with five frequencies per decade. Note that the error
    is clipped for values smaller than 0.01\,\% and bigger than 100\,\%.}
  \label{fig:GUI-FFTLog-lin}
\end{figure}
%

The example model is a marine scenario with 1\,km water depth of resistivity
$\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target of
100\,\ohmm at 1\,km below the seafloor in a background of 1\,\ohmm. The source
is 50\,m above the seafloor, the receiver is on the seafloor, and the response
is the inline $x$-directed $E$-field. The left subplot shows the imaginary part
in the frequency domain and the right plot the corresponding impulse response
in the time domain. The red lines are the \replaced{semi-analytical
responses}{precise results obtained from \empymod}. The blue circles indicate
the actually computed responses and the black dots the frequencies which are
interpolated or set to zero. The resulting time-domain response has a relative
error of less than 1\,\% everywhere except for very early times.

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\begin{figure*}
  \centering
  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02a-GUI-FFTLog-log}}%
  \label{fig:GUI-FFTLog-log}}\hfill
  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02b-GUI-DLF-lin}}%
  \label{fig:GUI-DLF-lin}}\\
  \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
  scale. (b) Using DLF instead of FFTLog, but computing the same
  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
  response has comparable accuracy (compare the error to the error in Figure
  \ref{fig:GUI-FFTLog-lin}).}
  \label{fig:GUI-log-lin}
\end{figure*}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. \added{But even so we only need to compute
50\,\% of the required frequencies in the case of the FFTLog, and only 14\,\%
in the case of the DLF.} \replaced{W}{However, w}ith this frequency-selection
we can model a wide range of offsets. This is shown in Figure
\ref{fig:multi-offset}, where the above parameters were used to model the
responses at offsets $r=1.5,3,6$, and 12\,km.
%
\begin{figure*}
  \centering
  \includegraphics[width=0.75\fwidth]{03-multi-offset}
  \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
    $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
    \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
    responses, the black dots are the responses which are set to zero or
    interpolated. The \replaced{grey}{black} curves are the analytical
    responses.}
  \label{fig:multi-offset}
\end{figure*}
%

The shortest offset defines the highest required frequency, and the largest
offset the lowest required frequency. So the more one can restrict the
necessary offset range, the \replaced{fewer}{less} frequencies are needed.
Another important factor is how to interpolate and extrapolate from the
computed frequencies to the frequencies required for the Fourier transform. For
the FFTLog only extrapolation for higher and lower frequencies is required. The
EM response becomes highly oscillatory for high frequencies, which makes it
very hard to extrapolate the response to frequencies $f>f_\mathrm{max}$.
However, if $f_\mathrm{max}$ is chosen judiciously, the importance of higher
frequencies for the Fourier transform can be neglected and we can set those
responses to zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be
changed to an interpolation by assuming a zero imaginary response at zero
frequency, and we then use PCHIP to interpolate the missing frequencies (as we
work on a logarithmic scale we cannot choose 0\,Hz \added{nor 0\,V/m}, but
instead take $10^{-100}$\,Hz \added{and $10^{-100}$\,V/m}). In the case of the
DLF method we also have to interpolate in between the computed frequencies, for
which we found it better to use a cubic spline. As can be seen from Figures
\ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF
with the same actually computed frequencies results in very similar responses.

@@

The main difference with \cite{GEO.08.Mulder} is that their adaptive gridding
searches for the optimal stretching factor $\alpha$ fulfilling certain
criteria, for a fixed number of cells. Our adaptive gridding, on the other
hand, searches for the smallest number of cells that still fulfil the given
criteria. The number of cells becomes therefore also a function of frequency.
\added{It is important to note that this is our implementation of an adaptive
grid, but there are certainly other possibilities. The relevant point for fast
computations is that the adaptive gridding tries to minimize the number of
required cells while being fine enough and at the same time large enough to
give accurate results. This is generally best done in a frequency and
conductivity dependent manner.} To go from the model grid to the computational
grid, we use the volume-averaging technique on logarithmic resistivities, as
used in \cite{GEO.07.Plessix}. While this technique ensures that the total
resistivity in the subsurface remains the same, it does not consider
effective-medium theory \citep{GEO.03.Davydycheva}, for instance, the apparent
anisotropy from a stack of finely layered formations of varying resistivity.
